---
title: "Take Home Final Part 2"
author: "Saumya Jain, Lauren Nelson"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I. Measures of Central Tendency

1. Mean

GDP Per Capita (PPP)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
mean <- mean(MyData$GDP_Per_Capita)
print(mean)
```

Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
mean <- mean(MyData$Male_Life_Expectancy)
print(mean)
```

2. Median

GDP Per Capita (PPP)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
median <- median(MyData$GDP_Per_Capita)
print(median)
```

Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
median <- median(MyData$Male_Life_Expectancy)
print(median)
```

3. Mode

GDP Per Capita (PPP)

No Mode

Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
mode <- getmode(MyData$Male_Life_Expectancy)
print(mode)
```

II. Measures of Variation
1. Range

GDP Per Capita (PPP)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
range <- range(MyData$GDP_Per_Capita, na.rm = FALSE)
range1 <- (range[2] - range[1])
print(range1)
```

Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
range <- range(MyData$Male_Life_Expectancy, na.rm = FALSE)
range1 <- (range[2] - range[1])
print(range1)
```

2. Variance

GDP Per Capita (PPP)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
variance <- var(MyData$GDP_Per_Capita, na.rm = FALSE)
print(variance)
```

Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
variance <- var(MyData$Male_Life_Expectancy, na.rm = FALSE)
print(variance)
```

3. Standard Deviation

GDP Per Capita (PPP)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
sd <- sd(MyData$GDP_Per_Capita, na.rm = FALSE)
print(sd)
```

Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
sd <- sd(MyData$Male_Life_Expectancy, na.rm = FALSE)
print(sd)
```

III. Correlation
GDP Per Capita (PPP) vs. Male Life Expectancy (Years)

```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
cor.test(Data$GDP_Per_Capita, Data$Male_Life_Expectancy)
```

IV. Scatterplot & Boxplots

GDP Per Capita (PPP) vs. Male Life Expectancy (Years) Scatterplot

```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
plot(Data$GDP_Per_Capita, Data$Male_Life_Expectancy,main="GDP Per Capita (PPP) vs. Male Life Expectancy (Years)",xlab="GDP per Capita", ylab="Life Expectancy")
```

GDP Per Capita (PPP) Boxplot
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
boxplot(Data$GDP_Per_Capita)
```

Male Life Expectancy (Years) Boxplot
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
boxplot(Data$Male_Life_Expectancy)
```

V. Normality Graph with Regression Line

GDP Per Capita (PPP)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
qqnorm(Data$GDP_Per_Capita)
qqline(Data$GDP_Per_Capita)
```

Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
qqnorm(Data$Male_Life_Expectancy)
qqline(Data$Male_Life_Expectancy)
```


VI. Outlier Check

GDP Per Capita(PPP)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
FindOutliers <- function(data) {
  lowerq = quantile(data)[2]
  upperq = quantile(data)[3]
  iqr = upperq - lowerq
  extreme.threshold.upper = (iqr * 1.5) + upperq
  extreme.threshold.lower = lowerq - (iqr * 1.5)
  result <- which(data > extreme.threshold.upper | data < extreme.threshold.lower)
}

temp <- FindOutliers(Data$GDP_Per_Capita)
testData <- Data[-temp]


```
Outliers: 88848, 135050 (2)


Male Life Expectancy (Years)
```{r}
setwd("C:/Users/Sam/Desktop")
MyData <- read.csv("Data.csv", header = TRUE, sep = ",")
Data <- c(MyData)
FindOutliers <- function(data) {
  lowerq = quantile(data)[2]
  upperq = quantile(data)[3]
  iqr = upperq - lowerq
  extreme.threshold.upper = (iqr * 1.5) + upperq
  extreme.threshold.lower = lowerq - (iqr * 1.5)
  result <- which(data > extreme.threshold.upper | data < extreme.threshold.lower)
}

temp <- FindOutliers(Data$Male_Life_Expectancy)
testData <- Data[-temp]
```
Outliers: None